R Notebook

### 2017-05-09 clean storm data.Rmd
R code for Poisson and NB posterior inference

---
title: "R Notebook"
output: html_notebook
---
```{r warning = FALSE, message = FALSE, echo = FALSE}
setwd("/users/nikopaulson/documents/homework/georgetown/math640")
library(tidyverse)

@@

>>>>

---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
```{r warning = FALSE, message = FALSE, echo = FALSE}
setwd("/users/nikopaulson/documents/homework/georgetown/math640")
library(tidyverse)

@@

```{r echo = FALSE}
summaryData2 <- df %>%
  filter(EVENT_TYPE == "TORNADO") %>%
  group_by(YEAR) %>%
  summarize(TOR_COUNT = n(), DEATHS = sum(DEATHS_DIRECT)) 
```

#POISSON

Likelihood:      $\mathcal{L}(X|\lambda)\propto\lambda^{n\bar{x}}e^{-n\lambda}$

Jeffreys' Prior: $\pi(\lambda)=\lambda^{1/2-1}e^{-0\cdot\lambda}$

Posterior:       $p(\lambda|X)=\lambda^{n\bar{x}+1/2-1}e^{-n\lambda}$

So the posterior on lambda with non-informative prior is $\mathcal{G}amma(n\bar{x}+1/2,n)$.

```{r}
set.seed(05092017)

df_tor <- df %>%
  filter(EVENT_TYPE == "TORNADO")

X <- df_tor$DEATHS_DIRECT
alpha <- sum(X) + 1 / 2
beta <- length(X)
B <- 20000

post_lambda <- rgamma(B, alpha, beta)
ci_lambda <- quantile(post_lambda, c(0.025, 0.975))
```

Using the data based off of 20000 posterior samples, we get the posterior median for $\lambda$ is `r round(median(post_lambda), 4)` with a 95% credible interval of `r round(ci_lambda[1], 4)` to `r round(ci_lambda[2], 4)`. The posterior denisty is in the figure below:

```{r echo = FALSE}
plot(density(post_lambda),
     xlab = expression(lambda),
     ylab = expression(paste("p(", lambda, "|x)")),
     main = "Density plot with 95% credible interval", 
     lwd = 2,
     col = "blue")
credInt <- quantile(post_lambda, probs = c(0.025, 0.975)) 
densOut <- density(post_lambda)
idStart <- max(which(densOut$x < credInt[1])) + 1 
idEnd <- min(which(densOut$x > credInt[2])) - 1
gx <- densOut$x[idStart : idEnd]
gy <- densOut$y[idStart : idEnd]
px <- rep(0, length(gy))
polygon(c(gx, rev(gx)), c(px, rev(gy)), border = FALSE, col = rgb(0, 0, 1, alpha = 0.5)) 
abline(h = 0)
```

#NEGATIVE BINOMIAL

Likelihood: $\mathcal{L}(X|r,p)=\Bigg[\prod_{i=1}^n\frac{\Gamma(r+x_i)}{\Gamma(r)x_i!}\Bigg]p^{n\bar{x}}(1-p)^{nr}$

Jeffreys' Prior: $\pi(r,p)=r^{1/2}p^{-1}(1-2)^{-1/2}$

Full Posterior: $p(r,p|X)=\mathcal{L}(X|r,p)\pi(r,p)$

Conditional of p: $p(p|r,X)\propto p^{n\bar{x}-1}(1-p)^{nr+1/2-1}$

Conditional of r: $p(r|p,X)\propto\big[\prod_{i=1}^n\Gamma(r+x_i)\big]\Gamma(r)^{-n}(1-p)^{nr}r^{1/2}$

```{r}
set.seed(05092017)

X <- df_tor$DEATHS_DIRECT

r <- 1 
alpha <- sum(X)
beta <- length(X) * r + 1 / 2
B <- 20000

post_p <- rbeta(B, alpha, beta)
ci_p <- quantile(post_p, c(0.025, 0.975))
```

Using the data based off of 20000 posterior samples, we get the posterior median for $p$ is `r round(median(post_p), 4)` with a 95% credible interval of `r round(ci_p[1], 4)` to `r round(ci_p[2], 4)`. The posterior denisty is in the figure below:

```{r echo = FALSE}
plot(density(post_p),
     xlab = expression(p),
     ylab = expression(paste("p(", p, "|r,x)")),
     main = "Density plot with 95% credible interval", 
     lwd = 2,
     col = "blue")
credInt <- quantile(post_p, probs = c(0.025, 0.975)) 
densOut <- density(post_p)
idStart <- max(which(densOut$x < credInt[1])) + 1 
idEnd <- min(which(densOut$x > credInt[2])) - 1
gx <- densOut$x[idStart : idEnd]
gy <- densOut$y[idStart : idEnd]
px <- rep(0, length(gy))
polygon(c(gx, rev(gx)), c(px, rev(gy)), border = FALSE, col = rgb(0, 0, 1, alpha = 0.5)) 
abline(h = 0)
```
